---
execute:
  echo: false
  cache: false
date: today
format:
  html:
    code-fold: true
    code-tools: true
    code-summary: "Show code"
---

# Annual QA/QC Pipeline: 2025 Data {.unnumbered}

## Introduction

This document contains the complete data ingestion, QA/QC evaluation, flagging, and CDX export workflow for the 2025 Kenai River Baseline Water Quality Monitoring dataset.

The structure of this document follows the Data Evaluation Checklist provided by the Alaska Department of Environmental Conservation. Field observations that do not meet QA/QC standards are flagged before upload to the EPA Water Quality Exchange (WQX).

For the 2021 worked example (with full narrative and documentation), see the report repo: <https://github.com/Kenai-Watershed-Forum/kenai-river-wqx>

------------------------------------------------------------------------

## Year Configuration

```{r year-config, message=FALSE, warning=FALSE}
# ── YEAR CONFIGURATION ─────────────────────────────────────────────────────────
# This is the ONLY block that must be updated for each new data year.
# All downstream scripts read from the `cfg` list constructed below.


library(tidyverse)
library(readxl)
library(openxlsx)
library(data.table)
library(stringr)
library(magrittr)
library(janitor)
library(hms)
library(lubridate)
library(anytime)
library(leaflet)
library(here)

year <- 2025

# Field sampling dates
spring_sample_date <- "4/30/2025"
summer_sample_date <- "7/23/2025"

spring_data_dir <- here("other/input/2025/spring_2025_wqx_data")
summer_data_dir <- here("other/input/2025/summer_2025_wqx_data")
output_qaqc_dir <- here("other/output/intermediate", year)

# Final WQX/CDX upload products land directly in other/output/
wqx_intermediate_path <- here("other/output", paste0(year, "_kwf_baseline_results_wqx.csv"))
flagged_export_path   <- here("other/output", paste0(year, "_export_data_flagged.csv"))
flag_decisions_path   <- here("other/input/wqx_templates",
                               paste0(year, "_data_flag_decisions.csv"))

cfg <- list(
  year                  = year,
  templates_dir         = here("other/input/wqx_templates"),
  wqx_template_file     = here("other/input/wqx_templates",
                                paste0("AWQMS_KWF_Baseline_", year, ".xlsx")),
  spring_data_dir       = spring_data_dir,
  summer_data_dir       = summer_data_dir,
  output_qaqc_dir       = output_qaqc_dir,
  wqx_intermediate_path = wqx_intermediate_path,
  flagged_export_path   = flagged_export_path,
  flag_decisions_path   = flag_decisions_path,
  wqx_downloads_dir     = here("other/input/WQX_downloads"),
  spring_sample_date    = spring_sample_date,
  summer_sample_date    = summer_sample_date
)

dir.create(here("other/output"),                                                    recursive = TRUE, showWarnings = FALSE)
dir.create(output_qaqc_dir,                                                         recursive = TRUE, showWarnings = FALSE)
dir.create(here("other/output/intermediate/field_qa_qc_data/metals_total_diss"),    recursive = TRUE, showWarnings = FALSE)
dir.create(here("other/output/intermediate/field_qa_qc_data/completeness_measures"),recursive = TRUE, showWarnings = FALSE)
dir.create(here("other/output/intermediate/misc"),                                  recursive = TRUE, showWarnings = FALSE)
dir.create(here("other/output/intermediate/lab_qaqc_data"),                         recursive = TRUE, showWarnings = FALSE)
```

------------------------------------------------------------------------

## Part A: Data Ingestion

**2025-specific notes:**

- Spring sampling: April 30, 2025. Summer sampling: July 23, 2025.
- SGS EDD delivered as XLSX (Sheet8) for both seasons. No ALS dissolved metals supplement this year.
- No BTEX (8260D) analysis in spring 2025; BTEX analyzed at 4 sites in summer 2025.
- TSS from SWWTP: both seasons read from the `Updated_Formatting` sheet. Spring source is `KRWF TSS MONITORING 05-01-25.xlsx`; summer is `KRWF TSS MONITORING 07-25-25.xls` (skip = 1 past title row). The original spring `.xls` (wide block format) is retained in the same directory for reference.
- FC from SWWTP in standard XLS format (spring: skip = 11; summer: skip = 10). Summer time sampled stored as HHMM integer.
- YSI ProQuatro / Hach turbidimeter data: single file covering both seasons, read in below.

### SGS Lab Results

```{r part-a-sgs, message=FALSE, warning=FALSE}
# Lab QC sample type codes -- segregated from field results
lab_qc_types <- c("MB", "LCS", "LCSD", "MS", "MSD", "CB", "OS",
                  "ICV", "CCV", "LLQC", "QCS", "ICB", "CVC", "IB")

# ── Read SGS EDD from Sheet8 ────────────────────────────────────────────────────
sgs_raw <- read_excel(
  file.path(cfg$spring_data_dir, "SGS",
            "spring_2025_kenai_baseline_sgs_results.xlsx"),
  sheet = "Sheet8"
) |>
  clean_names() |>
  remove_empty()

# ── Parse date/time fields (stored as "MM/DD/YYYY HH:MM" character strings) ────
sgs_raw <- sgs_raw |>
  mutate(
    collect_dt   = mdy_hm(collect_date),
    collect_date = as.Date(collect_dt),
    collect_time = as_hms(collect_dt),
    rec_dt       = mdy_hm(rec_date),
    rec_date     = as.Date(rec_dt),
    rec_time     = as_hms(rec_dt),
    run_dt       = mdy_hm(run_date_time),
    run_date     = as.Date(run_dt),
    run_time     = as_hms(run_dt)
  ) |>
  select(-collect_dt, -rec_dt, -run_dt, -run_date_time)

# ── Strip "EP" prefix from analytical_method (EP200.8 -> 200.8) ────────────────
sgs_raw <- sgs_raw |>
  mutate(analytical_method = str_replace(analytical_method, "^EP", ""))

# ── Assign result_sample_fraction from DISSOLVED flag ──────────────────────────
# DISSOLVED == "L" = lab-filtered (dissolved fraction)
# DISSOLVED == "." = unfiltered (total/recoverable)
# Nutrients (SM21 methods): fraction handled downstream by format_wqx.R lookup table
sgs_raw <- sgs_raw |>
  mutate(
    result_sample_fraction = case_when(
      dissolved == "L" & analytical_method == "200.8" ~ "Dissolved",
      dissolved == "." & analytical_method == "200.8" ~ "Unfiltered",
      TRUE ~ NA_character_
    )
  )

# ── Join analysis code matching table (SGS codes -> EPA method IDs) ─────────────
analysis_codes <- read_excel(
  file.path(cfg$templates_dir, "analysis_code_matching_table.xlsx")
) |>
  clean_names() |>
  select(sgs_analysis_code, epa_analysis_id, context_code) |>
  mutate(sgs_analysis_code = str_replace(sgs_analysis_code, "^EP", "")) |>
  rename(analytical_method = sgs_analysis_code)

sgs_raw <- left_join(sgs_raw, analysis_codes, by = "analytical_method")

# ── Join site names matching table -> monitoring_location_id + sample_condition ──
# sample_condition and monitoring_location_id both come from the spreadsheet;
# no string-based derivation needed.
sgs_site_names <- read_excel(
  file.path(cfg$templates_dir, "sgs_site_names_matching_table.xlsx")
) |>
  clean_names() |>
  select(sgs_sample_id, monitoring_location_id, sample_condition) |>
  mutate(sgs_sample_id = str_trim(sgs_sample_id)) |>
  rename(sample_id = sgs_sample_id)

sgs_raw <- sgs_raw |>
  mutate(sample_id = str_trim(sample_id)) |>
  left_join(sgs_site_names, by = "sample_id")

# ── Segregate lab QC rows to output file ────────────────────────────────────────
sgs_qaqc <- sgs_raw |>
  filter(sample_type %in% lab_qc_types)

dir.create(cfg$output_qaqc_dir, recursive = TRUE, showWarnings = FALSE)
write.csv(sgs_qaqc,
          file.path(cfg$output_qaqc_dir, "sgs_als_qaqc_dat.csv"),
          row.names = FALSE)
cat("Lab QC rows written:", nrow(sgs_qaqc), "\n")

# ── Filter to field results; rename to dat contract ─────────────────────────────
dat_sgs <- sgs_raw |>
  filter(!sample_type %in% lab_qc_types, !is.na(result)) |>
  mutate(
    lab_name   = "SGS",
    result     = as.numeric(result),
    result_detection_condition = case_when(
      grepl("U", resultflag) ~ "Not Detected",
      TRUE ~ NA_character_
    ),
    note = sdg
  ) |>
  select(
    sample              = sample_id,
    lab_sample          = lab_sample_id,
    lab_name,
    analyte,
    result,
    units,
    resultflag,
    loq,
    lod,
    collect_date,
    collect_time,
    rec_date,
    rec_time,
    run_date,
    run_time,
    analytical_method,
    epa_analysis_id,
    context_code,
    monitoring_location_id,
    note,
    sample_type,
    sample_condition,
    dissolved,
    matrix,
    result_sample_fraction,
    result_detection_condition
  )

dat_sgs_spring <- dat_sgs

# ── Summer 2025 SGS ──────────────────────────────────────────────────────────────
sgs_raw_sum <- read_excel(
  file.path(cfg$summer_data_dir, "SGS",
            "summer_2025_kenai_baseline_results_sgs.xlsx"),
  sheet = "Sheet8"
) |>
  clean_names() |>
  remove_empty()

sgs_raw_sum <- sgs_raw_sum |>
  mutate(
    collect_dt   = mdy_hm(collect_date),
    collect_date = as.Date(collect_dt),
    collect_time = as_hms(collect_dt),
    rec_dt       = mdy_hm(rec_date),
    rec_date     = as.Date(rec_dt),
    rec_time     = as_hms(rec_dt),
    run_dt       = mdy_hm(run_date_time),
    run_date     = as.Date(run_dt),
    run_time     = as_hms(run_dt)
  ) |>
  select(-collect_dt, -rec_dt, -run_dt, -run_date_time)

sgs_raw_sum <- sgs_raw_sum |>
  mutate(analytical_method = str_replace(analytical_method, "^EP", ""))

sgs_raw_sum <- sgs_raw_sum |>
  mutate(
    result_sample_fraction = case_when(
      dissolved == "L" & analytical_method == "200.8" ~ "Dissolved",
      dissolved == "." & analytical_method == "200.8" ~ "Unfiltered",
      TRUE ~ NA_character_
    )
  )

sgs_raw_sum <- left_join(sgs_raw_sum, analysis_codes, by = "analytical_method")

sgs_raw_sum <- sgs_raw_sum |>
  mutate(sample_id = str_trim(sample_id)) |>
  left_join(sgs_site_names, by = "sample_id")

sgs_qaqc_sum <- sgs_raw_sum |>
  filter(sample_type %in% lab_qc_types)
write.csv(sgs_qaqc_sum,
          file.path(cfg$output_qaqc_dir, "sgs_als_qaqc_dat_summer.csv"),
          row.names = FALSE)
cat("Summer lab QC rows written:", nrow(sgs_qaqc_sum), "\n")

dat_sgs_summer <- sgs_raw_sum |>
  filter(!sample_type %in% lab_qc_types, !is.na(result)) |>
  mutate(
    lab_name   = "SGS",
    result     = as.numeric(result),
    result_detection_condition = case_when(
      grepl("U", resultflag) ~ "Not Detected",
      TRUE ~ NA_character_
    ),
    note = sdg
  ) |>
  select(
    sample              = sample_id,
    lab_sample          = lab_sample_id,
    lab_name,
    analyte,
    result,
    units,
    resultflag,
    loq,
    lod,
    collect_date,
    collect_time,
    rec_date,
    rec_time,
    run_date,
    run_time,
    analytical_method,
    epa_analysis_id,
    context_code,
    monitoring_location_id,
    note,
    sample_type,
    sample_condition,
    dissolved,
    matrix,
    result_sample_fraction,
    result_detection_condition
  )

dat_sgs <- bind_rows(dat_sgs_spring, dat_sgs_summer)

# ── Expand trip blanks to both covered sites ─────────────────────────────────
# Trip blanks labelled "RMx&y trip blank" span two monitoring locations.
# Each is duplicated so both sites receive a WQX activity row.
trip_blank_sites <- tribble(
  ~sample,                 ~monitoring_location_id,
  "RM1.5&6.5 trip blank",  10000002L,   # RM 1.5 – Kenai City Dock
  "RM1.5&6.5 trip blank",  10000005L,   # RM 6.5 – Cunningham Park
  "RM40&43 trip blank",    10000027L,   # RM 40  – Bing's Landing
  "RM40&43 trip blank",    10000028L    # RM 43  – Upstream of Dow Island
)

dat_sgs <- bind_rows(
  dat_sgs |> filter(!is.na(monitoring_location_id)),
  dat_sgs |>
    filter(is.na(monitoring_location_id)) |>
    select(-monitoring_location_id) |>
    left_join(trip_blank_sites, by = "sample")
)

cat("SGS field rows:", nrow(dat_sgs),
    " (spring:", nrow(dat_sgs_spring), "| summer:", nrow(dat_sgs_summer),
    "| trip blank expansion adds", nrow(trip_blank_sites) / 2, "x 9 rows)\n")
cat("SGS sites:", n_distinct(dat_sgs$monitoring_location_id), "\n")
cat("SGS analytes:", n_distinct(dat_sgs$analyte), "\n")
```

### Fecal Coliform Results

```{r part-a-fc, message=FALSE, warning=FALSE}
fc_path_spring <- file.path(cfg$spring_data_dir, "SWWTP", "KRWF Fecal 04-30-25.xls")

# SWWTP XLS: skip 11 rows to reach column headers
# Columns: Dish Number | Sample Location/RM | ML | Time Sampled | Time In |
#          Time Out | Colony Count | Colony Count/100mL
fc_raw <- read_excel(fc_path_spring, skip = 11, col_types = "text")

# FC site -> monitoring_location_id lookup (inline; same sites as SGS)
fc_site_lkp <- tribble(
  ~site_rm,        ~monitoring_location_id, ~sample_condition,
  "RM 0",          10000008L,               NA_character_,
  "RM 1.5",        10000002L,               NA_character_,
  "RM 6.5",        10000005L,               NA_character_,
  "RM 10",         10000015L,               NA_character_,
  "RM 10.1",       10000016L,               NA_character_,
  "RM 12.5",       10000017L,               NA_character_,
  "RM 18",         10000018L,               NA_character_,
  "RM 18 DUP",     10000018L,               "Field Duplicate",
  "RM 19",         10000020L,               NA_character_,
  "RM 21",         10000021L,               NA_character_,
  "RM 22",         10000022L,               NA_character_,
  "RM 22 DUP",     10000022L,               "Field Duplicate",
  "RM 23",         10000023L,               NA_character_,
  "RM 30",         10000024L,               NA_character_,
  "RM 31",         10000025L,               NA_character_,
  "RM 36",         10000026L,               NA_character_,
  "RM 40",         10000027L,               NA_character_,
  "RM 43",         10000028L,               NA_character_,
  "RM 44",         10000029L,               NA_character_,
  "RM 50",         10000030L,               NA_character_,
  "RM 70",         10000031L,               NA_character_,
  "RM 74",         10000032L,               NA_character_,
  "RM 79.5",       10000424L,               NA_character_,
  "RM 82",         10000425L,               NA_character_
)

dat_fc <- fc_raw |>
  rename(
    site_rm          = `Sample Location/RM`,
    time_sampled     = `Time Sampled`,
    result_per_100ml = `Colony Count/100mL`,
    dish_number      = `Dish Number`
  ) |>
  # Remove blanks, positives, and sites with N/A results
  filter(
    !is.na(site_rm),
    !grepl("^BLANK|^POSITIVE", site_rm, ignore.case = TRUE),
    result_per_100ml != "N/A",
    !is.na(result_per_100ml)
  ) |>
  mutate(
    result       = as.numeric(result_per_100ml),
    # Time Sampled is stored as an Excel decimal fraction of a day
    collect_time = as_hms(as.numeric(time_sampled) * 86400),
    collect_date = mdy(spring_sample_date),
    rec_date     = mdy(spring_sample_date),
    rec_time     = as_hms(0),
    run_date     = mdy("5/1/2025"),  # SWWTP analysis date from file name
    run_time     = as_hms(0),
    resultflag   = case_when(result == 0 ~ "U", TRUE ~ "="),
    result_detection_condition = case_when(
      result == 0 ~ "Not Detected", TRUE ~ NA_character_
    ),
    analyte             = "Fecal Coliform",
    units               = "MPN/100ml",
    analytical_method   = "9222D",
    epa_analysis_id     = "9222D",
    context_code        = "APHA",
    lab_name            = "SWWTP",
    lab_sample          = dish_number,
    sample_type         = "PS",
    matrix              = "Water",
    dissolved           = NA_character_,
    note                = NA_character_,
    loq                 = NA_real_,
    lod                 = NA_real_,
    result_sample_fraction = NA_character_
  ) |>
  left_join(fc_site_lkp, by = "site_rm") |>
  filter(!is.na(monitoring_location_id)) |>
  select(
    sample              = site_rm,
    lab_sample,
    lab_name,
    analyte,
    result,
    units,
    resultflag,
    loq,
    lod,
    collect_date,
    collect_time,
    rec_date,
    rec_time,
    run_date,
    run_time,
    analytical_method,
    epa_analysis_id,
    context_code,
    monitoring_location_id,
    note,
    sample_type,
    sample_condition,
    dissolved,
    matrix,
    result_sample_fraction,
    result_detection_condition
  )

dat_fc_spring <- dat_fc

# ── Summer 2025 FC ───────────────────────────────────────────────────────────────
# Summer DUP sites differ from spring: RM 10 DUP and RM 23 DUP
fc_site_lkp_sum <- tribble(
  ~site_rm,        ~monitoring_location_id, ~sample_condition,
  "RM 0",          10000008L,               NA_character_,
  "RM 1.5",        10000002L,               NA_character_,
  "RM 6.5",        10000005L,               NA_character_,
  "RM 10",         10000015L,               NA_character_,
  "RM 10 DUP",     10000015L,               "Field Duplicate",
  "RM 10.1",       10000016L,               NA_character_,
  "RM 12.5",       10000017L,               NA_character_,
  "RM 18",         10000018L,               NA_character_,
  "RM 19",         10000020L,               NA_character_,
  "RM 21",         10000021L,               NA_character_,
  "RM 22",         10000022L,               NA_character_,
  "RM 23",         10000023L,               NA_character_,
  "RM 23 DUP",     10000023L,               "Field Duplicate",
  "RM 30",         10000024L,               NA_character_,
  "RM 31",         10000025L,               NA_character_,
  "RM 36",         10000026L,               NA_character_,
  "RM 40",         10000027L,               NA_character_,
  "RM 43",         10000028L,               NA_character_,
  "RM 44",         10000029L,               NA_character_,
  "RM 50",         10000030L,               NA_character_,
  "RM 70",         10000031L,               NA_character_,
  "RM 74",         10000032L,               NA_character_,
  "RM 79.5",       10000424L,               NA_character_,
  "RM 82",         10000425L,               NA_character_
)

fc_path_summer <- file.path(cfg$summer_data_dir, "SWWTP", "KRWF Fecal 07-23-25.xls")

# Summer FC: skip=10 (one fewer header row than spring). Time Sampled stored as
# 4-digit HHMM integer (e.g. 1057 = 10:57), not Excel decimal fraction.
fc_raw_sum <- read_excel(fc_path_summer, skip = 10, col_types = "text")

dat_fc_summer <- fc_raw_sum |>
  rename(
    site_rm          = `Sample Location/RM`,
    time_sampled     = `Time Sampled`,
    result_per_100ml = `Colony Count/100mL`,
    dish_number      = `Dish Number`
  ) |>
  filter(
    !is.na(site_rm),
    !grepl("^BLANK|^POSITIVE", site_rm, ignore.case = TRUE),
    result_per_100ml != "N/A",
    !is.na(result_per_100ml)
  ) |>
  mutate(
    result = as.numeric(result_per_100ml),
    # Time Sampled in summer is HHMM integer (e.g. 1057 = 10:57)
    collect_time = {
      t <- as.numeric(time_sampled)
      as_hms(floor(t / 100) * 3600 + (t %% 100) * 60)
    },
    collect_date = mdy(summer_sample_date),
    rec_date     = mdy(summer_sample_date),
    rec_time     = as_hms(0),
    run_date     = mdy("7/24/2025"),  # next-day analysis; verify against SWWTP lab report
    run_time     = as_hms(0),
    resultflag   = case_when(result == 0 ~ "U", TRUE ~ "="),
    result_detection_condition = case_when(
      result == 0 ~ "Not Detected", TRUE ~ NA_character_
    ),
    analyte             = "Fecal Coliform",
    units               = "MPN/100ml",
    analytical_method   = "9222D",
    epa_analysis_id     = "9222D",
    context_code        = "APHA",
    lab_name            = "SWWTP",
    lab_sample          = dish_number,
    sample_type         = "PS",
    matrix              = "Water",
    dissolved           = NA_character_,
    note                = NA_character_,
    loq                 = NA_real_,
    lod                 = NA_real_,
    result_sample_fraction = NA_character_
  ) |>
  left_join(fc_site_lkp_sum, by = "site_rm") |>
  filter(!is.na(monitoring_location_id)) |>
  select(
    sample              = site_rm,
    lab_sample,
    lab_name,
    analyte,
    result,
    units,
    resultflag,
    loq,
    lod,
    collect_date,
    collect_time,
    rec_date,
    rec_time,
    run_date,
    run_time,
    analytical_method,
    epa_analysis_id,
    context_code,
    monitoring_location_id,
    note,
    sample_type,
    sample_condition,
    dissolved,
    matrix,
    result_sample_fraction,
    result_detection_condition
  )

dat_fc <- bind_rows(dat_fc_spring, dat_fc_summer)

cat("FC rows:", nrow(dat_fc),
    " (spring:", nrow(dat_fc_spring), "| summer:", nrow(dat_fc_summer), ")\n")
```

### Total Suspended Solids

```{r part-a-tss, message=FALSE, warning=FALSE}
# Spring TSS: read from the tidy Updated_Formatting sheet in the new XLSX file.
# Site names in this file retain original SWWTP quirks ("RM O", "RM22", "RM22 DUP"),
# so the lookup table below matches them exactly without further normalization.
tss_path_spring <- file.path(cfg$spring_data_dir, "SWWTP",
                              "KRWF TSS MONITORING 05-01-25.xlsx")

tss_raw_spr <- read_excel(tss_path_spring, sheet = "Updated_Formatting") |>
  filter(sample_type == "PS") |>
  mutate(
    # Sample_Time stored as Excel time-only (date = 1899-12-31 artifact)
    collect_time = as_hms(format(Sample_Time, "%H:%M:%S")),
    result       = as.numeric(`S.S.mg/L`)
  )

# TSS site name -> monitoring_location_id
# Matches original SWWTP site name quirks: "RM O" (capital O), "RM22" (no space)
tss_site_lkp <- tribble(
  ~site,       ~monitoring_location_id, ~sample_condition,
  "RM O",      10000008L,               NA_character_,
  "RM 1.5",    10000002L,               NA_character_,
  "RM 6.5",    10000005L,               NA_character_,
  "RM 10",     10000015L,               NA_character_,
  "RM 10.1",   10000016L,               NA_character_,
  "RM 12.5",   10000017L,               NA_character_,
  "RM 18",     10000018L,               NA_character_,
  "RM 18 DUP", 10000018L,               "Field Duplicate",
  "RM 19",     10000020L,               NA_character_,
  "RM 21",     10000021L,               NA_character_,
  "RM22",      10000022L,               NA_character_,
  "RM22 DUP",  10000022L,               "Field Duplicate",
  "RM 23",     10000023L,               NA_character_,
  "RM 30",     10000024L,               NA_character_,
  "RM 31",     10000025L,               NA_character_,
  "RM 36",     10000026L,               NA_character_,
  "RM 40",     10000027L,               NA_character_,
  "RM 43",     10000028L,               NA_character_,
  "RM 44",     10000029L,               NA_character_,
  "RM 50",     10000030L,               NA_character_,
  "RM 70",     10000031L,               NA_character_,
  "RM 74",     10000032L,               NA_character_,
  "RM 79.5",   10000424L,               NA_character_,
  "RM 82",     10000425L,               NA_character_
)

# ADAPT: Verify LOD and LOQ against current QAPP and SWWTP lab report.
tss_lod <- 0.5   # mg/L placeholder
tss_loq <- 1.0   # mg/L placeholder

dat_tss <- tss_raw_spr |>
  left_join(tss_site_lkp, by = c("Sample_Location" = "site")) |>
  filter(!is.na(monitoring_location_id)) |>
  mutate(
    collect_date    = mdy(spring_sample_date),
    rec_date        = mdy("5/1/2025"),
    rec_time        = as_hms(0),
    run_date        = mdy("5/1/2025"),
    run_time        = as_hms(0),
    analyte         = "Total suspended solids",
    units           = "mg/l",
    lod             = tss_lod,
    loq             = tss_loq,
    resultflag = case_when(
      result < tss_lod ~ "U",
      result < tss_loq ~ "J",
      TRUE ~ "="
    ),
    result = case_when(result < tss_lod ~ NA_real_, TRUE ~ result),
    result_detection_condition = case_when(
      resultflag == "U" ~ "Not Detected", TRUE ~ NA_character_
    ),
    analytical_method      = "2540-D",
    epa_analysis_id        = "2540-D",
    context_code           = "APHA",
    lab_name               = "SWWTP",
    lab_sample             = as.character(row_number()),
    sample_type            = "PS",
    matrix                 = "Water",
    dissolved              = NA_character_,
    note                   = NA_character_,
    result_sample_fraction = NA_character_
  ) |>
  select(
    sample              = Sample_Location,
    lab_sample,
    lab_name,
    analyte,
    result,
    units,
    resultflag,
    loq,
    lod,
    collect_date,
    collect_time,
    rec_date,
    rec_time,
    run_date,
    run_time,
    analytical_method,
    epa_analysis_id,
    context_code,
    monitoring_location_id,
    note,
    sample_type,
    sample_condition,
    dissolved,
    matrix,
    result_sample_fraction,
    result_detection_condition
  )

dat_tss_spring <- dat_tss

# ── Summer 2025 TSS ──────────────────────────────────────────────────────────────
# Summer TSS file has a tidy "Updated_Formatting" sheet (skip=1 past title row).
# Site names use underscores: "RM_0", "RM_10_DUP", etc.
tss_path_summer <- file.path(cfg$summer_data_dir, "SWWTP",
                              "KRWF TSS MONITORING 07-25-25.xls")

tss_raw_sum <- read_excel(tss_path_summer, sheet = "Updated_Formatting", skip = 1) |>
  filter(sample_type == "PS") |>
  mutate(
    # Normalize underscores and DUP variants to consistent "RM X [DUP]" format
    site_clean = Sample_Location |>
      str_replace_all("_DUP$",  " DUP") |>
      str_replace_all("_",      " ") |>
      str_replace("(?i) DUP$", " DUP") |>
      str_squish(),
    sample_condition = if_else(
      str_detect(str_to_lower(Sample_Location), "dup"),
      "Field Duplicate", NA_character_
    ),
    # Sample_Time is stored as Excel time-only (date = 1899-12-31 artifact)
    collect_time = as_hms(format(Sample_Time, "%H:%M:%S")),
    result       = as.numeric(`S.S.mg/L`)
  )

# Re-use the spring tss_site_lkp but with summer site name variants
tss_site_lkp_sum <- tribble(
  ~site_clean,     ~monitoring_location_id,
  "RM 0",          10000008L,
  "RM 1.5",        10000002L,
  "RM 6.5",        10000005L,
  "RM 10",         10000015L,
  "RM 10 DUP",     10000015L,
  "RM 10.1",       10000016L,
  "RM 12.5",       10000017L,
  "RM 18",         10000018L,
  "RM 19",         10000020L,
  "RM 21",         10000021L,
  "RM 22",         10000022L,
  "RM 23",         10000023L,
  "RM 23 DUP",     10000023L,
  "RM 30",         10000024L,
  "RM 31",         10000025L,
  "RM 36",         10000026L,
  "RM 40",         10000027L,
  "RM 43",         10000028L,
  "RM 44",         10000029L,
  "RM 50",         10000030L,
  "RM 70",         10000031L,
  "RM 74",         10000032L,
  "RM 79.5",       10000424L,
  "RM 82",         10000425L
)

dat_tss_summer <- tss_raw_sum |>
  left_join(tss_site_lkp_sum, by = "site_clean") |>
  filter(!is.na(monitoring_location_id), !is.na(result)) |>
  mutate(
    collect_date    = mdy(summer_sample_date),
    rec_date        = mdy("7/25/2025"),  # from file name: KRWF TSS MONITORING 07-25-25
    rec_time        = as_hms(0),
    run_date        = mdy("7/25/2025"),
    run_time        = as_hms(0),
    analyte         = "Total suspended solids",
    units           = "mg/l",
    lod             = tss_lod,
    loq             = tss_loq,
    resultflag = case_when(
      result < tss_lod ~ "U",
      result < tss_loq ~ "J",
      TRUE ~ "="
    ),
    result = case_when(result < tss_lod ~ NA_real_, TRUE ~ result),
    result_detection_condition = case_when(
      resultflag == "U" ~ "Not Detected", TRUE ~ NA_character_
    ),
    analytical_method      = "2540-D",
    epa_analysis_id        = "2540-D",
    context_code           = "APHA",
    lab_name               = "SWWTP",
    lab_sample             = as.character(row_number()),
    sample_type            = "PS",
    matrix                 = "Water",
    dissolved              = NA_character_,
    note                   = NA_character_,
    result_sample_fraction = NA_character_
  ) |>
  select(
    sample              = site_clean,
    lab_sample,
    lab_name,
    analyte,
    result,
    units,
    resultflag,
    loq,
    lod,
    collect_date,
    collect_time,
    rec_date,
    rec_time,
    run_date,
    run_time,
    analytical_method,
    epa_analysis_id,
    context_code,
    monitoring_location_id,
    note,
    sample_type,
    sample_condition,
    dissolved,
    matrix,
    result_sample_fraction,
    result_detection_condition
  )

dat_tss <- bind_rows(dat_tss_spring, dat_tss_summer)

cat("TSS rows:", nrow(dat_tss),
    " (spring:", nrow(dat_tss_spring), "| summer:", nrow(dat_tss_summer), ")\n")
```

### YSI ProQuatro / Turbidity Field Measurements

```{r part-a-ysi, message=FALSE, warning=FALSE}
# Single file covers both seasons. Season is inferred from collection date.
# Each site has 2 replicate observations; these are averaged per site-date-parameter.
# DUP sites (e.g. "Soldotna Creek-DUP") are separated by the site lookup table.
# Physically impossible values excluded (negative pH, DO > 20 mg/L).

ysi_path <- here("other/input/2025",
                  "2025 Kenai Agency Baseline YSI ProQuatro and Turbidity Data.xlsx")

ysi_param_map <- c(
  "Temperature"  = "Water Temperature",
  "Conductivity" = "Specific Conductance",
  "DO"           = "Dissolved Oxygen",
  "Turbidity"    = "Turbidity",
  "pH"           = "pH"
)
ysi_method_map <- c(
  "Water Temperature"    = "170.1",
  "Specific Conductance" = "120.1",
  "Dissolved Oxygen"     = "360.1",
  "Turbidity"            = "180.1",
  "pH"                   = "150.1"
)
ysi_unit_map <- c(
  "Water Temperature"    = "deg C",
  "Specific Conductance" = "uS/cm",
  "Dissolved Oxygen"     = "mg/L",
  "Turbidity"            = "NTU",
  "pH"                   = "None"
)

# YSI site name -> monitoring_location_id lookup
ysi_site_lkp <- tribble(
  ~ysi_site_name,              ~monitoring_location_id, ~sample_condition,
  "No Name Creek",             10000008L,               NA_character_,
  "City of Kenai Dock",        10000002L,               NA_character_,
  "Kenai Dock",                10000002L,               NA_character_,
  "Cunningham Park",           10000005L,               NA_character_,
  "Beaver Creek",              10000015L,               NA_character_,
  "Beaver Creek- DUP",         10000015L,               "Field Duplicate",
  "Kenai River",               10000016L,               NA_character_,
  "Kenai River (RM10.1)",      10000016L,               NA_character_,
  "Pillars",                   10000017L,               NA_character_,
  "Pillars Boat Launch",       10000017L,               NA_character_,
  "Poachers Cove",             10000018L,               NA_character_,
  "Poachers Cove -DUP",        10000018L,               "Field Duplicate",
  "Slikok Creek",              10000020L,               NA_character_,
  "Soldotna Bridge",           10000021L,               NA_character_,
  "Soldotna Creek",            10000022L,               NA_character_,
  "Soldotna Creek-DUP",        10000022L,               "Field Duplicate",
  "Swiftwater Park",           10000023L,               NA_character_,
  "Swiftwater Park-DUP",       10000023L,               "Field Duplicate",
  "Funny River",               10000024L,               NA_character_,
  "Morgan's Landing",          10000025L,               NA_character_,
  "Morgans Landing",           10000025L,               NA_character_,
  "Moose River",               10000026L,               NA_character_,
  "Bing's Landing",            10000027L,               NA_character_,
  "Bings Landing",             10000027L,               NA_character_,
  "Upstream of Dow Island",    10000028L,               NA_character_,
  "Dow Island",                10000028L,               NA_character_,
  "Mouth of Killey River",     10000029L,               NA_character_,
  "Killey River",              10000029L,               NA_character_,
  "Skilak Lake Overflow",      10000030L,               NA_character_,
  "Jim's Landing",             10000031L,               NA_character_,
  "Jims Landing (RM 70)",      10000031L,               NA_character_,
  "Russian River",             10000032L,               NA_character_,
  "Russian River (RM74)",      10000032L,               NA_character_,
  "Juneau Creek",              10000424L,               NA_character_,
  "Juneau Creek (RM 79.5)",    10000424L,               NA_character_,
  "Kenai Lake Bridge",         10000425L,               NA_character_,
  "Kenai Lake (RM 82)",        10000425L,               NA_character_
)

ysi_all <- read_excel(ysi_path) |>
  filter(
    !is.na(`Site Name`), !is.na(Parameter), !is.na(Value),
    !(Parameter == "pH" & (Value < 0 | Value > 14)),
    !(Parameter == "DO"  & Value > 20)
  ) |>
  mutate(
    analyte      = ysi_param_map[Parameter],
    collect_date = as.Date(`Site Depart Date`),
    collect_time = as_hms(format(`Collection Time (HH:MM)`, "%H:%M:%S"))
  ) |>
  filter(!is.na(analyte))

# Join site lookup, then average replicates per site-date-parameter group
dat_ysi <- ysi_all |>
  left_join(ysi_site_lkp, by = c("Site Name" = "ysi_site_name")) |>
  filter(!is.na(monitoring_location_id)) |>
  group_by(monitoring_location_id, sample_condition, analyte, collect_date) |>
  summarise(
    result       = mean(Value, na.rm = TRUE),
    collect_time = first(collect_time),
    .groups      = "drop"
  ) |>
  mutate(
    sample                     = as.character(monitoring_location_id),
    lab_sample                 = NA_character_,
    lab_name                   = "YSI",
    units                      = ysi_unit_map[analyte],
    resultflag                 = "=",
    loq                        = NA_real_,
    lod                        = NA_real_,
    rec_date                   = collect_date,
    rec_time                   = collect_time,
    run_date                   = collect_date,
    run_time                   = collect_time,
    analytical_method          = ysi_method_map[analyte],
    epa_analysis_id            = ysi_method_map[analyte],
    context_code               = "USEPA",
    note                       = NA_character_,
    sample_type                = "PS",
    dissolved                  = NA_character_,
    matrix                     = "Water",
    result_sample_fraction     = "Total",
    result_detection_condition = NA_character_
  ) |>
  select(
    sample,
    lab_sample,
    lab_name,
    analyte,
    result,
    units,
    resultflag,
    loq,
    lod,
    collect_date,
    collect_time,
    rec_date,
    rec_time,
    run_date,
    run_time,
    analytical_method,
    epa_analysis_id,
    context_code,
    monitoring_location_id,
    note,
    sample_type,
    sample_condition,
    dissolved,
    matrix,
    result_sample_fraction,
    result_detection_condition
  )

cat("YSI rows:", nrow(dat_ysi), "\n")
cat("YSI sites:", n_distinct(dat_ysi$monitoring_location_id), "\n")
cat("YSI parameters:", paste(sort(unique(dat_ysi$analyte)), collapse = ", "), "\n")

unmapped_ysi <- ysi_all |>
  filter(!`Site Name` %in% ysi_site_lkp$ysi_site_name) |>
  pull(`Site Name`) |> unique()
if (length(unmapped_ysi) > 0) {
  cat("⚠ Unmapped YSI sites (excluded):", paste(unmapped_ysi, collapse = ", "), "\n")
}
```

### Bind All Results into `dat`

```{r part-a-bind, message=FALSE, warning=FALSE}
dat <- bind_rows(dat_sgs, dat_fc, dat_tss, dat_ysi) |>
  mutate(monitoring_location_id = as.character(monitoring_location_id))

cat("Total dat rows:", nrow(dat), "\n")
cat("Analytes:", paste(sort(unique(dat$analyte)), collapse = ", "), "\n")
cat("Sites:", n_distinct(dat$monitoring_location_id), "\n")
```

------------------------------------------------------------------------

## Part B: WQX Formatting

```{r format-wqx, message=FALSE, warning=FALSE, include=FALSE}
# Applies all lookup table joins (coordinates, result sample fractions, detection
# condition, preservative, container), constructs WQX-formatted column names,
# builds Activity IDs, and writes the provisional WQX intermediate CSV.
#
# NOTE: format_wqx.R was updated to handle pre-existing result_sample_fraction
# (set in Part A for EP200.8 dissolved/total distinction). Part A values take
# precedence; the lookup table fills in unset fractions (nutrients, FC, TSS).
#
# Requires: dat, cfg   Produces: dat, dat_raw, all_dat; writes cfg$wqx_intermediate_path
source(here("other/misc/qaqc_repo_transition/functions/format_wqx.R"))
```

### Provisional Results (Prior to QA/QC Review)

The following table summarizes results after WQX formatting, before QA/QC evaluation.

```{r provisional-summary, message=FALSE}
read.csv(cfg$wqx_intermediate_path) |>
  clean_names() |>
  count(characteristic_name, result_analytical_method_id, activity_start_date) |>
  arrange(activity_start_date, characteristic_name)
```

------------------------------------------------------------------------

## Part C: QA/QC Checklist

Prior to upload to the EPA WQX, all water quality data is checked against a standard Data Evaluation Checklist developed in coordination with the Alaska Department of Environmental Conservation.

#### Pre-Database

##### Overall Project Success

IN PROGRESS HERE 4/22/2026

**1.) Were the appropriate analytical methods used for all parameters?**

```{r q01}
# MANUAL: Confirm analytical methods match those in the approved QAPP.
# Spring 2025 methods: EP200.8 (metals), SM21 4500NO3-F (nitrate), SM21 4500P-B,E (TP),
# SM 9222D (fecal coliform), SM 2540-D (TSS).
```
Yes. Analytical methods from the approved 2023 QAPP were employed.

<br>

**2.) Were there any deviations from the sampling plan?**

```{r q02}
# MANUAL: Describe any deviations from the sampling plan (sites not visited,
# schedule changes, etc.). If no deviations, state so explicitly.
```

<br>

**3.) Were field duplicates, blanks, and/or other QC samples collected as planned?**

```{r q03-planned, message=FALSE, warning=FALSE}
# ADAPT: Create other/input/2025/spring_2025_wqx_data/planned_samples_2025.xlsx
# following the template format to enable full planned-vs-actual comparison.
# Until that file exists, actual sample counts are used as a proxy.

total_samples_collected_summary <- read.csv(cfg$wqx_intermediate_path) |>
  clean_names() |>
  mutate(activity_start_date = ymd(activity_start_date)) |>
  filter(activity_type %in% c("Field Msr/Obs",
                               "Quality Control Field Replicate Msr/Obs",
                               "Quality Control Sample-Trip Blank")) |>
  group_by(result_analytical_method_id, activity_start_date, activity_type) |>
  count() |>
  rename(actual_results_n = n)

# Use actual as planned for CMB denominator (conservative estimate)
planned_samples_summary <- total_samples_collected_summary |>
  rename(expected_results_n = actual_results_n)

planned_samples_summary_by_site <- read.csv(cfg$wqx_intermediate_path) |>
  clean_names() |>
  filter(activity_type == "Field Msr/Obs") |>
  group_by(monitoring_location_id, result_analytical_method_id,
           activity_start_date, activity_type) |>
  count() |>
  rename(expected_results_n = n)

total_samples_collected_summary
```

```{r q03-narrative}
# MANUAL: Describe any deviations between planned and actual results.
```

<br>

**4.) Do the laboratory reports provide results for all sites and parameters?**

```{r q04}
# MANUAL: Confirm based on Q3 results. Note any exceptions.
```

<br>

**5.) Is a copy of the Chain of Custody included with the laboratory reports?**

```{r q05}
# SGS COC on file: other/input/2025/spring_2025_wqx_data/SGS/1251731_COC.pdf
# SWWTP: FC and TSS results sheets serve as lab records.
# MANUAL: Confirm SWWTP documentation meets project requirements.
```

<br>

**6.) Do the laboratory reports match the Chain of Custody and requested methods throughout?**

```{r q06}
# MANUAL
```

<br>

**7.) Are the number of samples on the laboratory reports the same as on the Chain of Custody?**

```{r q07}
# MANUAL
```

<br>

**8.) Was all supporting info provided in the laboratory report, such as reporting limits for all analyses and definitions?**

```{r q08}
# MANUAL
```

<br>

**9.) Are site names, dates, and times correct and as expected?**

```{r q09, message=FALSE}
sites_in_data <- read.csv(cfg$wqx_intermediate_path) |>
  clean_names() |>
  distinct(monitoring_location_id, activity_start_date) |>
  arrange(monitoring_location_id, activity_start_date)
sites_in_data
# MANUAL: Confirm all expected sites present and dates correct.
```

<br>

**10.) Were there any issues with instrument calibration?**

```{r q10}
# MANUAL
```

<br>

**11.) Did the instruments perform as expected?**

```{r q11}
# MANUAL
```

<br>

**12.) Was instrument calibration performed according to the QAPP and instrument recommendations?**

```{r q12}
# MANUAL
```

<br>

**13.) Was instrument verification during the field season performed according to the QAPP and instrument recommendations?**

```{r q13}
# MANUAL
```

<br>

**14.) Were instrument calibration verification logs or records kept?**

```{r q14}
# MANUAL
```

<br>

**15.) Do the instrument data files site IDs, time stamps and file names match?**

```{r q15}
# MANUAL
```

<br>

**16.) Is any insitu field data rejected and why?**

```{r q16}
# MANUAL
```

<br>

**17.) Were preservation, hold time and temperature requirements met?**

```{r q17-holdtimes, message=FALSE, warning=FALSE}
col_names <- c("sample", "epa_analysis_id", "analyte", "collect_date", "collect_time",
               "rec_date", "rec_time", "lab_name")

holdtime_dat <- dat |>
  select(any_of(col_names)) |>
  mutate(
    rec_date1 = as.Date(rec_date),
    rec_time1 = case_when(
      !is.na(rec_time) ~ as_hms(rec_time),
      TRUE             ~ as_hms(0)
    )
  ) |>
  select(-rec_time, -rec_date) |>
  rename(rec_time = rec_time1, rec_date = rec_date1) |>
  mutate(
    activity_datetime          = ymd_hms(paste(collect_date, collect_time)),
    rec_datetime               = ymd_hms(paste(rec_date, rec_time)),
    hold_time_hours            = as.numeric(difftime(rec_datetime, activity_datetime,
                                                      units = "hours")),
    result_analytical_method_id = epa_analysis_id
  )

max_holding_times <- read.csv(
  file.path(cfg$templates_dir, "sample_holding_times.csv")
) |>
  transform(max_holding_time_hours = as.numeric(max_holding_time_hours))

holdtime_dat <- left_join(holdtime_dat, max_holding_times,
                          by = "result_analytical_method_id") |>
  mutate(hold_time_pass = case_when(
    hold_time_hours > max_holding_time_hours ~ "N",
    TRUE ~ "Y"
  ))

write.csv(holdtime_dat,
          here("other/output/intermediate/field_qa_qc_data/holding_time_calcs.csv"),
          row.names = FALSE)

cat("Hold time failures:", sum(holdtime_dat$hold_time_pass == "N", na.rm = TRUE), "\n")
holdtime_dat |> filter(hold_time_pass == "N")
```

<br>

**18.) Are dissolved metal quantities less than total metals quantities?**

```{r q18-dissolved-total, message=FALSE, warning=FALSE}
# Spring 2025: Calcium, Iron, Magnesium measured as total only.
# Copper and Zinc have both dissolved and total -- check those pairs.

diss_total_analytes <- c("Copper", "Zinc")

diss_total_dat <- dat |>
  filter(analyte %in% diss_total_analytes, !is.na(result),
         is.na(sample_condition)) |>  # primary samples only; exclude DUPs and blanks
  mutate(
    result_mg_l = case_when(units == "ug/L" ~ result / 1000, TRUE ~ result),
    fraction    = case_when(
      result_sample_fraction == "Dissolved"  ~ "Dissolved",
      TRUE                                   ~ "Total"
    )
  )

dir_diss <- here("other/output/intermediate/field_qa_qc_data/metals_total_diss")

for (elem in diss_total_analytes) {
  elem_dat <- diss_total_dat |>
    filter(analyte == elem) |>
    select(monitoring_location_id, collect_date, fraction, result_mg_l) |>
    pivot_wider(names_from = fraction, values_from = result_mg_l) |>
    mutate(
      total_greater_eq_diss = case_when(
        is.na(Total) | is.na(Dissolved) ~ NA,
        Total >= Dissolved              ~ "Y",
        TRUE                            ~ "N"
      )
    )
  write.csv(elem_dat, file.path(dir_diss, paste0(tolower(elem), ".csv")))
  cat(elem, "— rows with total >= dissolved:",
      sum(elem_dat$total_greater_eq_diss == "Y", na.rm = TRUE),
      "of", nrow(elem_dat), "\n")
}
```

```{r q18-narrative}
# MANUAL: Interpret results. Note any cases where dissolved > total.
```

<br>

**19.) Are the duplicate sample(s) RPD (Relative Percent Difference) within range described in QAPP?**

```{r q19-rpd, message=FALSE, warning=FALSE}
qaqc_sites <- read_excel(
  file.path(cfg$templates_dir, "wqx_qaqc/wqx_qaqc_info.xlsx"),
  sheet = paste0("field_dup_sites_", year)
) |> remove_empty()

export_dat_rpd <- read.csv(cfg$wqx_intermediate_path) |> clean_names()
field_dup_dat  <- inner_join(export_dat_rpd, qaqc_sites, by = "monitoring_location_id")

rpd_cols <- c("monitoring_location_id", "activity_type", "activity_start_date",
              "characteristic_name", "result_detection_condition",
              "result_value", "result_unit",
              "result_detection_limit_type_1", "result_detection_limit_value_1",
              "result_detection_limit_unit_1")

rpd_check_dat <- field_dup_dat |>
  select(any_of(rpd_cols)) |>
  mutate(result_value = as.numeric(result_value)) |>
  pivot_wider(names_from = "activity_type", values_from = "result_value",
              values_fn = \(x) x[1]) |>
  clean_names() |>
  mutate(
    rpd_eligible = !is.na(field_msr_obs) &
                   !is.na(quality_control_field_replicate_msr_obs),
    rpd_pct = abs(
      (field_msr_obs - quality_control_field_replicate_msr_obs) /
        ((field_msr_obs + quality_control_field_replicate_msr_obs) / 2) * 100
    )
  )

rpd_eligible_dat <- rpd_check_dat |>
  filter(rpd_eligible) |>
  mutate(threshold = case_when(
    grepl("Nitrate|Nitrite",  characteristic_name, ignore.case = TRUE) ~ 25,
    grepl("Phosphorus",       characteristic_name, ignore.case = TRUE) ~ 25,
    grepl("suspended|solids", characteristic_name, ignore.case = TRUE) ~ 5,
    TRUE ~ 20
  ))

rpd_exceed <- rpd_eligible_dat |> filter(!is.na(rpd_pct), rpd_pct > threshold)

write.csv(rpd_check_dat,
          here("other/output/intermediate/field_qa_qc_data/rpd_check_dat.csv"),
          row.names = FALSE)

cat("RPD-eligible pairs:", nrow(rpd_eligible_dat), "\n")
cat("Pairs exceeding QAPP threshold:", nrow(rpd_exceed), "\n")
if (nrow(rpd_exceed) > 0) rpd_exceed
```

```{r q19-narrative}
# MANUAL: Discuss RPD results. Below-LOQ pairs retained as Accepted non-detects.
```

<br>

**20.) Were there any laboratory discrepancies, errors, data qualifiers, or QC failures (review laboratory duplicates, matrix spikes and blanks)?**

```{r q20-matrix-spikes, message=FALSE, warning=FALSE}
ms_cols <- c("lab_name", "sample", "collect_date", "sample_type", "result",
             "analyte", "analytical_method", "resultflag", "percent_recovered",
             "rec_limit_low", "rec_limit_high", "sample_rpd",
             "rpd_limit_low", "rpd_limit_high", "loq", "lod",
             "detection_limit", "sample_condition")

matrix_spike_limits <- read.csv(
  file.path(cfg$output_qaqc_dir, "sgs_als_qaqc_dat.csv")
) |>
  select(any_of(ms_cols)) |>
  mutate(
    percent_recovered = as.numeric(percent_recovered),
    rec_limit_low     = as.numeric(rec_limit_low),
    rec_limit_high    = as.numeric(rec_limit_high),
    rec_limit_pass = case_when(
      is.na(rec_limit_low)                                              ~ "",
      percent_recovered > rec_limit_high |
        percent_recovered < rec_limit_low                              ~ "N",
      TRUE                                                              ~ "Y"
    )
  )

matrix_spike_fails <- matrix_spike_limits |> filter(rec_limit_pass == "N")

write.csv(matrix_spike_fails,
          here("other/output/intermediate/lab_qaqc_data",
               paste0("matrix_spike_recovery_fails_", year, ".csv")),
          row.names = FALSE)

cat("Matrix spike recovery failures:", nrow(matrix_spike_fails), "\n")
if (nrow(matrix_spike_fails) > 0) matrix_spike_fails
```

```{r q20-narrative}
# MANUAL: Describe additional lab QC anomalies not captured by matrix spike check.
```

<br>

**21.) Is any laboratory data rejected and why?**

```{r q21}
# MANUAL: Record flagging decisions in:
# other/input/wqx_templates/2025_data_flag_decisions.csv
# That file is joined in Part D (apply_qaqc_flags.R).
```

<br>

**22.) During the field season, review raw data files (EDDs, instrument records) as they are received. Document changes and corrections to methods as needed.**

```{r q22}
# MANUAL: Note any corrections applied during Part A ingestion and how documented.
# Spring 2025 known corrections:
#   - "RM O" (typo) in TSS file treated as "RM 0" (site 10000008).
#   - "RM22" (missing space) in TSS file treated as "RM 22" (site 10000022).
#   - SGS analytical_method "EP200.8" stripped to "200.8" for consistency.
```

<br>

**23.) Is the dataset complete and did you receive the expected number of results?**

```{r q23}
# MANUAL: Reference Q3. Note that summer 2025 data is not yet collected.
```

<br>

**24.) Was the data collected representative of environmental conditions?.**

```{r q24}
# MANUAL
```

<br>

**25.) Does project meet Completeness Measure A criteria?**

From the QAPP, CMA = primary samples collected / usable samples submitted (goal: 85%).

```{r apply-flags, message=FALSE, warning=FALSE}
# Join flag decisions and write flagged export CSV.
# IMPORTANT: This chunk must run before Q25 and Q26.
source(here("other/misc/qaqc_repo_transition/functions/apply_qaqc_flags.R"))
```

```{r q25-cma, message=FALSE}
analytes_qapp <- read.csv(
  file.path(cfg$templates_dir, "analytes_list_manual_edit.csv")
)
colnames(analytes_qapp) <- c("characteristic_name", "analyte_abbreviation")

completeness_wide <- read.csv(cfg$flagged_export_path) |>
  clean_names() |>
  group_by(result_analytical_method_id, characteristic_name, flag) |>
  count() |>
  rename(n_results = n) |>
  pivot_wider(names_from = "flag", values_from = "n_results")

# Y column absent when no results are flagged yet
if (!"Y" %in% names(completeness_wide)) completeness_wide$Y <- NA_integer_
if (!"N" %in% names(completeness_wide)) completeness_wide$N <- NA_integer_

completeness_summary_param <- completeness_wide |>
  rename(flag_Y = Y, flag_N = N) |>
  rowwise() |>
  mutate(
    total_samples = sum(flag_Y, flag_N, na.rm = TRUE),
    cma           = flag_N / total_samples,
    cma           = replace_na(cma, 0)
  )

write.csv(completeness_summary_param,
          here("other/output/intermediate/field_qa_qc_data/completeness_measures/cma_parameter.csv"))

cma_dat    <- read.csv(cfg$flagged_export_path) |> clean_names()
cma_total  <- nrow(cma_dat)
cma_n      <- sum(cma_dat$flag == "N", na.rm = TRUE)
cma_pct    <- paste0(round(cma_n / cma_total * 100, 1), "%")
cat("CMA (overall):", cma_pct, "(goal: 85%)\n")
completeness_summary_param
```

<br>

**26.) Does project meet Completeness Measure B criteria?**

From the QAPP, CMB = unflagged results / planned results (goal: 60%).

```{r q26-cmb, message=FALSE}
cmb_wide <- read.csv(cfg$flagged_export_path) |>
  clean_names() |>
  group_by(result_analytical_method_id, activity_start_date, activity_type, flag) |>
  count() |>
  pivot_wider(names_from = "flag", values_from = "n")

if (!"Y" %in% names(cmb_wide)) cmb_wide$Y <- NA_integer_
if (!"N" %in% names(cmb_wide)) cmb_wide$N <- NA_integer_

cmb_by_method <- cmb_wide |>
  rename(flag_Y = Y, flag_N = N) |>
  mutate(activity_start_date = ymd(activity_start_date)) |>
  left_join(planned_samples_summary,
            by = c("result_analytical_method_id", "activity_start_date",
                   "activity_type")) |>
  mutate(
    cmb = flag_N / expected_results_n,
    cmb = replace_na(cmb, 0)
  )

write.csv(cmb_by_method,
          here("other/output/intermediate/field_qa_qc_data/completeness_measures/cmb_parameter.csv"))

total_planned  <- sum(planned_samples_summary$expected_results_n, na.rm = TRUE)
cmb_n          <- sum(cmb_by_method$flag_N, na.rm = TRUE)
cmb_pct        <- paste0(round(cmb_n / total_planned * 100, 1), "%")
cat("CMB (overall):", cmb_pct, "(goal: 60%)\n")
cmb_by_method
```

<br>

**27.) Was the QA officer consulted for any data concerns?**

```{r q27}
# MANUAL
```

<br>

**28.) Are the correct monitoring locations associated with the project? Are latitude and longitude filled out in a consistent format?**

```{r q28-map, message=FALSE, warning=FALSE}
export_dat <- read.csv(cfg$flagged_export_path)

site_check <- export_dat |>
  distinct(monitoring_location_id, activity_latitude, activity_longitude) |>
  transform(
    activity_latitude  = as.numeric(activity_latitude),
    activity_longitude = as.numeric(iconv(activity_longitude, "utf-8", "ascii", sub = ""))
  ) |>
  mutate(activity_longitude = case_when(
    activity_longitude > 0 ~ activity_longitude * -1,
    TRUE ~ activity_longitude
  ))

if (knitr::is_html_output()) {
  leaflet(site_check) |>
    addTiles() |>
    addMarkers(~activity_longitude, ~activity_latitude,
               popup = ~as.character(monitoring_location_id),
               label  = ~as.character(monitoring_location_id))
} else {
  knitr::kable(site_check, caption = "2025 monitoring locations")
}
```

<br>

**29.) Are the QAPP and other supporting documents attached?**

```{r q29}
# MANUAL: Confirm QAPP filename is set correctly in generate_cdx_export.R.
```

<br>

**30.) Is all project metadata correct?**

```{r q30}
# MANUAL: Confirm organization ID, project ID, QAPP approval status, site coordinates.
```

<br>

**31.) Is the organization ID correct?**

```{r q31}
cat("Expected organization ID: Kenai_WQX\n")
```

<br>

**32.) Are the time zones consistent and correct (AKDT in summer)?**

```{r q32-timezone}
# Spring 2025 sampling (April 30) is in AKDT (Alaska Daylight Time).
# DST began March 9, 2025 -- all 2025 spring data is AKDT.
export_dat <- read.csv(cfg$flagged_export_path) |>
  mutate(time_zone = "AKDT")
write.csv(export_dat, cfg$flagged_export_path)
cat("Time zone applied: AKDT\n")
```

<br>

**33.) Are all media types included? Media types appropriate to Characteristic?**

```{r q33}
cat("Activity media types in dataset:\n")
unique(export_dat$activity_media_name)
```

<br>

**34.) Check Sample Collection, Preparation and Preservation Methods, Thermal Preservative, Equipment ID, Activity Media. Is supporting information included and correct?**

```{r q34}
cat("Result sample fractions:\n"); unique(export_dat$result_sample_fraction)
cat("\nChemical preservatives:\n"); unique(export_dat$chemical_preservative_used)
cat("\nContainer types:\n");        unique(export_dat$sample_container_type)
```

<br>

**35.) Are all expected activity types present and are QC samples correctly identified?**

```{r q35}
cat("Activity types:\n")
unique(export_dat$activity_type)
```

<br>

**36.) Is the Activity media subdivision filled in (if relevant)?**

```{r q36}
unique(export_dat$activity_media_subdivision_name)
```

<br>

**37.) For Water activity media, is the relative depth filled in?**

```{r q37}
unique(export_dat$activity_depth_height_measure)
# All surface grab samples recorded at ~15 cm depth.
```

<br>

**38.) Is the number of results for each Characteristic correct?**

```{r q38}
chr_num <- export_dat |>
  group_by(characteristic_name) |>
  tally()

write.csv(chr_num,
          here("other/output/intermediate/misc",
               paste0("characteristic_count_", year, ".csv")),
          row.names = FALSE)
chr_num
```

<br>

**39.) Do the range of result values make sense?**

```{r q39}
export_dat |>
  filter(!is.na(result_value)) |>
  group_by(characteristic_name, result_unit) |>
  summarise(min = min(result_value, na.rm = TRUE),
            max = max(result_value, na.rm = TRUE),
            n   = n(), .groups = "drop") |>
  arrange(characteristic_name)
# MANUAL: Compare against prior years and regulatory limits.
```

<br>

**40.) Are units correct and consistent for each parameter?**

```{r q40}
z <- export_dat |> distinct(characteristic_name, result_unit)
cat("Unique unit values:\n"); print(unique(z$result_unit))

# Standardize case
export_dat[export_dat == "mg/l"] <- "mg/L"
write.csv(export_dat, cfg$flagged_export_path)
```

<br>

**41.) Are detection limits and laboratory qualifiers included for analytical results?**

```{r q41}
cat("Result qualifiers:\n");     unique(export_dat$result_qualifier)
cat("\nDetection limit types:\n"); unique(export_dat$result_detection_limit_type_1)
```

<br>

**42.) Are results trip blanks and/or field blanks above detection limits, and is any related data flagged as a result?**

```{r q42}
trip_blanks <- export_dat |>
  filter(activity_type == "Quality Control Sample-Trip Blank")

cat("Trip blank results:", nrow(trip_blanks), "\n")
cat("Detections in trip blanks (result_value > 0):",
    sum(!is.na(trip_blanks$result_value) & trip_blanks$result_value > 0), "\n")

# MANUAL: If any trip blank analytes are detected, determine whether associated
# field results should be flagged and update flag_decisions_path accordingly.
```

------------------------------------------------------------------------

## Part D: Flag + CDX Export

```{r generate-cdx, message=FALSE, warning=FALSE}
# Apply Result Status ID (Accepted/Rejected), finalize column format, and write
# three CDX upload files: results_activities.csv, project.csv, station.csv.
#
# REVIEW: generate_cdx_export.R contains a filter that excludes Phosphorus from
# EPA Method 200.8 (not in the 2021 QAPP). Confirm whether this filter applies
# to 2025 before running. Comment out in that script if not applicable.
#
# Requires: export_dat (with time_zone applied in Q32), cfg
# Requires: WQP downloads in cfg$wqx_downloads_dir (populate manually before running)

wqp_project_file <- file.path(cfg$wqx_downloads_dir, "project/project.csv")

if (!file.exists(wqp_project_file)) {
  message("Part D skipped: WQP downloads not yet present in ", cfg$wqx_downloads_dir,
          "\nDownload project, station, and results files from WQP before running.")
} else {
  source(here("other/misc/qaqc_repo_transition/functions/generate_cdx_export.R"))
}
```

### CDX Upload Files

The following files are ready for upload to the EPA Water Quality Exchange Central Data Exchange (CDX) at <https://cdx.epa.gov/>:

```{r cdx-file-summary, message=FALSE}
cdx_files <- c(
  here("other/output/results_activities.csv"),
  here("other/output/project.csv"),
  here("other/output/station.csv")
)

for (f in cdx_files) {
  if (file.exists(f)) {
    d <- read.csv(f)
    cat(basename(f), ":", nrow(d), "rows\n")
  } else {
    cat(basename(f), ": NOT FOUND -- run generate_cdx_export.R\n")
  }
}
```
